Efficient Data Provenance Analysis for Concurrent Programs

### 2013-11-13 paper.tex
committing with 10 pages, redone related.  Need to add bugs?

{\bf TODO: Fix this figure to reflect availability of LWS info if the program crashes/is stopped at *any* point in the exec (like the read X in thread 2).}

@@

There are several categories of prior work related to last writer slices and
\ctraps.

\paragraph{Program Slicing}
Slicing techniques~\cite{tipslicingsurvey} are techniques for identifying
useful subsequences of a program's instructions.   Slicing can work by
analyzing program code statically or analyzing execution traces dynamically.
Many slicing techniques discussed in ~\cite{tipslicingsurvey} monitor control
and data dependences to identify which parts of a program might be relevant to
a task ({\em e.g., debugging}).

Thin Slicing~\cite{thinslicing} is especially related to this work. Thin
Slicing monitors and reports a select subset of dependent operations from an
execution.  The goal is to eliminate some operations that would be included in
a traditional slice to make the slice simpler and less costly to collect.
While this work and Thin Slicing differ in what they collect and how, our
motivation to monitor a select subset of dependences was similar to theirs.

%is unlike our work in that our work focused on
%monitoring inter-thread dependences in production code without sampling or
%hardware support.

\paragraph{Debugging with Data Provenance Information}

Prior work has looked at using data provenance information in
debugging~\cite{whylineicse, whylinechi}.  The WhyLine debugger allowed
programmers to ask questions about why particular variables held particular
values during debugging.  WhyLine answered those questions by presenting data-
and control-flows that influenced those variables.  WhyLine showed that
provenance information is intuitive to programmers and makes debugging easier,
justifying our focus on provenance information in our last writer slicing
design.

There are several characteristics of this work that distinguish it from prior
work on debugging with data provenance information.  Unlike prior work, our
work focuses on isolating a limited subset of dependence information that we
can collect with run time overheads low enough for production use.
Furthermore, we focus on {\em concurrent} program debugging.  Prior work on
provenance debugging that focused mainly on single threaded programs and had
dependence slicing overheads are probably too high for production.

Bad Apple Tracking~\cite{badapples} is also a form of debugging with data
provenance information. This work records the {\em origin} of an unusable value
and can report it to a programmer when it is used in a way that causes a
failure.

@@

CCI~\cite{cci}, Bugaboo~\cite{bugaboo}, Recon~\cite{recon}, and
DefUse~\cite{defuse} all explicitly track inter-thread memory dependences at
runtime.  Unlike our work, several of these techniques were not designed to run
in production systems~\cite{recon,defuse}.  A distinguishing characteristic of
our work is that it is effective and fast enough for production without the
need to aggregate information from multiple different sampled executions, like
CCI~\cite{cci}.  Additionally, our system does not require invasive hardware
changes like Bugaboo~\cite{bugaboo} or LBA~\cite{paralog}.

Other techniques~\cite{threadclustering,threadcriticality} use  hardware debug
registers to monitor dependences efficiently enough for production use.  These
techniques are similar to our work in that they track how threads interact.
These techniques are different from our work in that they track information
specific to the performance and our system provides general support.

\paragraph{Extensible Program Instrumentation}
There are many systems for building analyses using extensible program
instrumentation.  Binary instrumentation
systems~\cite{pin,dynamorio,valgrind,roadrunner} provide completely general
support to instrument arbitrary code -- instrumenting a program to track
communciation is possible in such systems.

These systems are similar to our work in that they aim to enable dynamic
analyses and runtime systems.  They are different from our work because they
rely on dynamic rewriting, yielding overheads often too high for deployment.
Instead, we trade generality for performance. Last writer slices and \ctraps
provide extensible support in the restricted domain of tracking communication
with overheads low enough for production.  Another difference is that our work
handles some of the delicate, performance sensitive code required to track
communication.  In general instrumentation systems, programmers write that code
from scratch at their own risk.

>>>>

@@

\paragraph{Debugging with Data Provenance Information}
The most related prior work looked at using various forms of provenance information for debugging.

We have discussed bad value origin tracking ~\cite{badapples} throughout this
work and compared directly with that technique in our evaluation.  Bad value
origin tracking is like our work because they store provenance information to
be used for debugging -- a similar goal to ours.  Bad value origin tracking
differs from our work in several ways.  First, their technique only tracks the
provenance of {\em unusable} values -- a fraction of all values in an
execution.  Our work, instead, tracks provenance for any value stored in a
potentially shared memory location, which includes a broader set of provenance
and targets our technique to concurrency debugging.  Second, their technique
does not run with overheads tolerable for performance for C and C++ programs.
They implement their technique for C and C++ programs using Valgrind MemCheck,
which has high runtime overheads.  Our technique has overheads low enough for
production in many cases, especially in server applications.  Third, their
technique does not explicity record thread information with tracked origins.
Our work focuses on concurrency and we include thread information.  As we show
in Section~\ref{sec:eval:debugging} thread information is important for
understanding some bugs.  Fourth, their technique uses bad value piggy-backing
to store origins, we use a shadow memory.

Like our work, the WhyLine debugger~\cite{whylineicse, whylinechi} allows
programmers to ask questions about why variables hold particular values during
debugging.  WhyLine answers by presenting data- and control-flows that
influenced those variables.  WhyLine differs from our work in that it uses
dependence slicing on recorded executions to answer provenance questions and
has overheads too high for production use.  WhyLine also did not explicitly
focus on concurrency.

\paragraph{Program Slicing}
Program slicing techniques identify useful subsequences of a program's
instructions.   Slicing can work by analyzing program code statically or
analyzing execution traces dynamically.  Many slicing techniques discussed in
the a recent survey~\cite{tipslicingsurvey} monitor control and data
dependences to identify which parts of a program might be relevant to a task
({\em e.g., debugging}).

Thin Slicing~\cite{thinslicing}, is a slicing technique that is related to our
work.  Thin Slicing yields the set of statements that computed on or wrote
values that influenced a particular variable's value (the ``seed'').  Thin
Slicing is similar in that its slices provide provenance and it excludes
control information like our work.  Thin Slicing is different in that it has
overheads too high for production and focused on minimizing the size 
of each slice.

@@

CCI~\cite{cci}, Bugaboo~\cite{bugaboo}, LBA~\cite{paralog}, Recon~\cite{recon},
and DefUse~\cite{defuse} all explicitly track some thread interactions.  Unlike
our work, several of these techniques were not designed to run in production
systems~\cite{recon,defuse}.  A distinguishing characteristic of our work is
that it is effective and fast enough for production without the need to
aggregate information from multiple different sampled executions, like
CCI~\cite{cci} does.  Additionally, our system does not require invasive
hardware changes like Bugaboo~\cite{bugaboo} or LBA~\cite{paralog}.

\paragraph{Extensible Program Instrumentation} 
There are many systems for building analyses using program instrumentation.
Binary instrumentation~\cite{pin,dynamorio,valgrind,roadrunner} is general and
tracking provenance and communciation is possible in such systems.

These systems are similar to \ctraps in that they enable dynamic analyses.
They are different from our work because they rely on dynamic binary
translation, yielding overheads often too high for deployment.  We trade some
generality for performance. \ctraps is effective mainly in the restricted
domain of concurrency analyses, but has overheads low enough for production.
Another difference is that our work handles some of the delicate, performance
sensitive code required concurrency analyses, like last writer tracking.  In
general instrumentation systems, programmers write that code from scratch at
their own risk.
